---
title: 'Reproducible Research: Peer Assessment 1'
output:
  html_document:
    keep_md: yes
  pdf_document: default
---

Initially setup our working directory and load the dplyr library.

```{r echo=TRUE, message=FALSE}
library(dplyr)

setwd("D:/dev/Coursera/assignments/rr_project1/RepData_PeerAssessment1")

```

## Loading and preprocessing the data

The activity file is loaded and processed.

```{r echo=TRUE}
activity_data = read.csv("data/activity.csv")
steps_per_day <- group_by(activity_data, date)
```


## What is mean total number of steps taken per day?

The table below shows the total, mean and median per day.

```{r echo=TRUE}
steps_summary <- summarize(steps_per_day, mean_per_day = mean(steps, na.rm=TRUE), median_per_day = median(steps,na.rm=TRUE), total_per_day = sum(steps))
print(select(steps_summary,date,mean_per_day,median_per_day,total_per_day),n=100)
hist(steps_summary$total_per_day)
```

## What is the average daily activity pattern?

The average daily activity pattern is as follows.

```{r}
steps_per_interval <- group_by(activity_data, interval)
interval_avg <- summarize (steps_per_interval, y=mean(steps, na.rm = TRUE))
plot(interval_avg$interval,interval_avg$y, type="l")
arrange(interval_avg,desc(y))
```

835 is the interval with the highest average number of steps across all days.

## Imputing missing values

1. The total number of rows with NAs is as follows...

```{r}
sum(is.na(activity_data))

```

2. We are going to populate the missing values by giving them the mean of that days intervals

```{r}
means <- summarize(steps_per_day, avg = mean(steps,na.rm = TRUE))
means[is.na(means)] = 0
new_activity_data <- activity_data
new_activity_data <- merge(new_activity_data,means)
new_activity_data <- mutate(new_activity_data, steps = ifelse(is.na(steps),avg,steps))

steps_per_day <- group_by(new_activity_data, date)
steps_summary <- summarize(steps_per_day, mean_per_day = mean(steps, na.rm=TRUE), median_per_day = median(steps,na.rm=TRUE), total_per_day = sum(steps))
print(select(steps_summary,date,total_per_day),n=100)
hist(steps_summary$total_per_day)

```

There is no real impact of inputting the missing values to the histogram or table. 

## Are there differences in activity patterns between weekdays and weekends?

There is more activity at the weekend.

```{r}

activity_data <- mutate(activity_data, day = as.factor(
                                        ifelse(weekdays(as.Date(date))=='Saturday'| weekdays(as.Date(date)) == 'Sunday',
                                        'Weekend','Weekday')))

weekday_steps <- subset(activity_data,day == 'Weekday')
weekend_steps <- subset(activity_data,day == 'Weekend')

steps_per_weekday <- group_by(weekday_steps, interval)
steps_per_weekend <- group_by(weekend_steps, interval)

weekday_summary <- summarize (steps_per_weekday, y=mean(steps, na.rm = TRUE))
weekend_summary <- summarize (steps_per_weekend, y=mean(steps, na.rm = TRUE))

par(mfrow = c(2, 1), mar = c(5, 4, 2, 1)) 
plot(weekday_summary$interval,weekday_summary$y, type="l")
plot(weekend_summary$interval,weekend_summary$y, type="l")

```
